---
title: "Reich - Tutorial 2"
author: "J Reich"
date: "1/17/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(survival)
```
 

```{r}
cougars <- read.csv("JAE_00481_Cougar Survival Data.csv")
head(cougars)
```

# Exercise 1

This data would fall under "interval censorship" type, since we know the start of data collection began above some time period (adulthood; although we don't have the exact data for each individual) and the end of the data collection came before another time period (death, in most cases).

# Exercise 2

Make a new dataset removing all truncated values:

```{r}
cougarsD <- cougars[cougars$Event == 1,]
```

### a. Calculate mean from new data set and compare to previous data set.

```{r}
mean(cougarsD$TotalMonth)
mean(cougars$TotalMonth)
```

The means are almost identical.

### b. Estimate the survival function of this new data set and compare to the first one we made:

```{r}
## original data graph:
survV <- Surv(time=cougars$TotalMonth, event=rep(1, nrow(cougars)))
fixV <- survfit(survV ~ 1)
plot(fixV)

## with new analysis in red:
survD <- Surv(time=cougarsD$TotalMonth)
fixD <- survfit(survD ~ 1)
points(fixD$surv~ fixD$time, type= "s", col="red")
```

We can see that the survival functions are quite similar to each other.

## c. 

```{r}

## Plot original truncated data
fixlm <- lm(TotalMonth ~ Housing.Density + Sex, data = cougars)
plot(TotalMonth~Housing.Density, data=cougars, pch=19, cex=0.5)
abline(fixlm$coef[1:2])

## Superimpose new analysis with removed data
regcougarsD <- lm(TotalMonth ~ Housing.Density + Sex, data = cougarsD)
points(TotalMonth~Housing.Density, data=cougarsD, pch=19, cex=0.5, col="red")
abline(regcougarsD$coef[1:2], col="red")


## compare Sex between two methods
plot(TotalMonth~Sex, data=cougars, pch=19, cex=0.5)
abline(fixlm$coef[c(1,3)])

points(TotalMonth~Sex, data=cougarsD, pch=19, cex=0.5, col = "red")
abline(regcougarsD$coef[c(1,3)], col = "red")

## Compare output between regression methods
summary(fixlm)
summary(regcougarsD)
```

Both methods of analysis have similar values for intercepts and coefficients, however, as expected, by deleting cases entirely, we end up with larger standard errors.

# Exercise 3. Demonstrate the difference between Kaplan-Meier and empirical survival:

```{r}
### Plotting the Kaplan-Meier survival function (given in Tutorial)
survKM <- Surv(time=cougars$TotalMonth, event=cougars$Event)
kmV <- survfit(survKM~1)
plot(kmV)

### Plotting the empirical survival function from before (red lines):
esf <- data.frame(matrix(ncol=2, nrow=length(unique(cougars$TotalMonth))))
colnames(esf) <- c("y", "s")
esf$y <- sort(unique(cougars$TotalMonth))

for(i in 1:nrow(esf)){
        esf$s[i] <- sum(cougars$TotalMonth>esf$y[i])/nrow(cougars)
}

points(esf, type="s", col="red")
```

The empirical survival function (red line) indicates an overall faster mortality rate than the Kaplan-Meier survival function. This makes sense, since the Kaplan-Meier estimate accounts for the censored data - that is, not every animal really died at the end of the study, and so we see that reflected in the way the curve is shifted to the right (as if "extra months" were added on to each cougar's life)


# Exercise 5

```{r}
deers <- read.csv("Fawn_survival_70day.csv")
```
```{r}
head(deers)
```
### a. Estimate the mean survival length of the fawns.

Calculate the total number of days each deer was monitored (an approximation of their age), and add the new column to the data set:

```{r}
deers$total.days <- (deers$k - deers$i)
head(deers)
```

Using the Kaplan-Meier estimate for interval censorship:

```{r}
survDInt <- Surv(time = deers$total.days, event = deers$fate)
DInt <- survfit(survDInt~1)
plot(DInt)
```

Estimate the mean survival lenght of fawns:
```{r}
print(DInt, print.rmean=TRUE)
```

So according to these estimates, the mean survival length of fawns is about 41 days.

## c. Investigate whether survival of fawns is affected by sex or timber harvesting.

```{r}
### run a tobit model on the survival curve we estimated above
deersreg <- survreg(survDInt~sex+timber, data=deers, dist = "gaussian")
summary(deersreg)
```

It appears that neither sex nor timber harvesting are significant factors in fawn survival. To see what this looks like graphically, examine some plots:

```{r}
boxplot(k~sex, data = deers)
abline(deersreg$coef[1:2])
```

```{r}
boxplot(k~timber, data = deers)
abline(deersreg$coef[c(1,3)])
```


. . .  these results don't make any sense :/